---
title: "Joseph Priestly's \"A Chart of Biography\""
author: "Jeffrey B. Arnold"
date: "2/5/2018"
output: html_document
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Joseph Priestly's "A Chart of Biography"}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```

## Introduction

Joseph Priestley's *A Chart of Biography*, first published in 1764, is the, or one of the, first instances of a modern timeline. (ADD Citation)

The **priestley** package is a data-only package containing datasets used in the 
the biographic information used in the  1764 and 1778 editions of the the *Chart of Biography*.


The data timelines include approximately two thousand individuals, "the most distinguished in the annals of fame".

> My rule, I considered, was renown and not merit; acquired fame,
> and not deserved reputation : so that a person who had
> made a great noise in the world, though he were
> known by nothing but the devastation he had made
> in it, was more acceptable to me than one who had
> deserved ever so well of it, if he had passed through
> it without being much known.

The timeline shows the names and lines indicating lifespan.
The individuals were grouped into five divisions:

The *Chart* was accompanied by a *A Description of a Chart of Biography: With a Catalouge of all the Names Inserted in It, and the Dates annexed to Them.*.
This included an index of the names, and 
more detailed information about the birth and death dates of the individuals included in the *Chart*,
as well as their profession.

about two thousand names of persons 

IT will easily occur to all my readers, that my
greatest difficulty must have been the proper choice of
names to fill this tablet of fame : and some degree of

solicitude is certainly unavoidable when a man volun-
tarily assumes the province of the arbiter and dispenser

of every man’s reputation, and when he sees all the
dead pass, as it were, in review before him for that
purpose. But this is no greater presumption than is
implied in numberless other works, and is in fact no

more than one man’s giving his present opinion con-
cerning others. The nature of the design necessarily

fixed some limits to the width of the divisions, and
though in some ages there was room enough for all
the candidates for fame ; if I would have inserted
them ; in others, and particularly in modern times,
where no reasonable space would admit a tenth part
of the candidates, it must require no small judgement
to decide concerning their respective pretensions.

With respect to this, I can only say that I have acquitted myself with all 
the impartiality of which I

was capable.HOWEVER a due regard to the nature of my design
made made the task much easier, and less invidious than

may, at first sight, be imagined. My rule, I con-
sidered, was renown and not merit ; acquired fame,

and not deserved reputation : so that a person who had
made a great noise in the world, though he were
known by nothing but the devastation he had made
in it, was more acceptable to me than one who had
deserved ever so well of it, if he had passed through
it without being much known. And it is manifest
that, if this chart be designed for the use of persons
who are beginning an acquaintance with history, it
must be drawn up according to this rule. Otherwise
it would not show them the situation of the names
they are most likely to look for in it.

I AM sensible it would be extreme arrogance in any

person to pretend to so accurate and extensive a know-
ledge even of the proportion of renown which all men

of every profession have acquired, as would enable him
to select the most distinguished names, with so much
exactness, that none of those who are omitted should
be more noted than any of those who are inserted :
yet thus much I hope the diligence of my inquiries
may encourage me to say, that, though some instances
may occur, to more diligent and longer continued
researches, of persons of greater note omitted, and
less note inserted, yet that the fame of all those

who are omitted would bear an exceeding small pro-
portion to the fame of those who are inserted.

I AM convinced, however, it is absolutely impos-
sible to please many persons with the choice of names

upon such a plan as this. No two persons living
would make the same choice. I will even venture
to say, from my own experience, that no one person
would, at different times, make the same choice. It
is only such an experience as I have had that can
convince any person, how extremely fluctuating are

our ideas of the relative fame of many different pers-
ons, and of the figure they have made in the world.

The many times I have altered my lists convinces me
that I should never revise them without seeing some
reason to make farther alterations ; but the many
times that I have replaced the same names after

having rejected them, convinces me that further al-
terations would have been of very little consequence.

BEFORE a person express great surprize at my o-
mitting or over-looking any favourite of his, let him

consider, particularly, at what time he flourished ;
for hundreds which have been excluded in later, and

more crowded ages, would have found a ready recep-
tion in an earlier period ; so that their exclusion must

not be attributed either to my fault or their own.
Let it always be remembered, that the celebrity of

any person is to be compared with that of his co-
temporaries, and not with that of those who made

their appearance either before or after him.

I HAVE perhaps inserted rather more Englishmen
in proportion, than those of other nations, but this

was not owing to any national prejudice ; but as re-
nown is a relative thing, regard must necessarily be

had to the minds of those into whose hands the chart
is most likely to fall. Had I drawn the chart in
order to its being published in any other part of the
world, even in France, I should naturally, with the
same ideas of relative fame, have made a different

choice of names. If the chart, as it stands at pre-
sent, seemed calculated rather for men of letters than

men of business, let it be attributed to a similar reason.

NEITHER the number, nor the width of the divi-
sions was determined at random, and a priori, but

both were made such as the names I previously
collected appeared most natural to require. Fewer
classes I could not have made, and at the same time
have preserved any tolerable distinctness, and more
would

THE classing of the names occasioned some little

difficulty to me, and may occasion some to the peru-
ser. My general rule has been to place every person

in that class in which he was most eminent. Thus

I have placed Mohammed, as the head of a celebra-
ted sect in religion, among the Divines ; though the

Caliphs, his successors, will be found among the
Statesmen.

The original chart is "about three feet in length,
and two feet in breadth" (Description A3)

```{r message = FALSE}
ggplot(priestley::Specimen) +
  geom_segment(aes(y = row, yend = row, x = start_2, xend = end_2),
               alpha = 0.5, size = 0.5) +
  geom_segment(aes(y = row, yend = row, x = start_1, xend = end_1),
               size = 1) +
  geom_text(aes(x = (start_2 + end_2) * 0.5, y = row, label = name),
            size = 3, vjust = "bottom", nudge_y = 0.25, color = "gray50") +
  facet_wrap(~ division, ncol = 1) +
  scale_x_continuous("", 
                     minor_breaks = seq(-600, 0, 50),
                     breaks = seq(-600, 0, 100)) +
  scale_y_continuous("", expand = c(0.1, 0.1)) +
  ggtitle("A Specimen of the Chart of Biography") +
  theme_minimal() +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank(),
        strip.background = element_rect(fill = "white", colour = NA),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank())
```


Wikimedia Commons contributors, "File:PriestleyChart.gif," Wikimedia Commons, the free media repository, <https://commons.wikimedia.org/w/index.php?title=File:PriestleyChart.gif&oldid=177413180> (accessed February 8, 2018).

## Descriptions

Examples in which the date of death and the age at death are known. These are drawn with a single solid black line since there is no uncertainty.

- Pindar d. 435 BC. 86 P
- Sophocles d. 406 BC 91 P
- Xenophon d. 359 BC. ab. 90 H
- Plato d. 348 BC. 81 HP
- Terence d. 159 BC. 35 P

> It is an imperfection which must necessarily attend every Chart of this
> nature, that the time of the death, but more especially the time of the birth
> of eminent men cannot always be found. In this case the compiler must content
> himself with placing his line as near as he can conjecture from history where
> his true place was, leaving marks to express the uncertainty
> there is attending it. The method I have used in
> this Chart is to express certainty by a full line, and
> what is uncertain by dots or a broken line, dispro
> of the dots in the following manner, according to the
> kind or degree of the uncertainty they have to express.
>
> If it be thought that a person was born or died but
> a little before or after a certain time, the full line
> begins or ends at the certain time, and a single dot only
> is placed at the beginning or end of the line

Examples include,

- Herodotus b. 484 BC. d. af. 413 H

> If history informs us that he was born or died
> about a certain time, a dot is placed just under the
> beginning or end of the full line

Examples include,

- Agis b. 24. BC. ab. 22
- Thucydides d. ab. 391. ab. 80 H

> If the birth be certain but the death uncertain, the line begins full but ends in dots.

Examples include,

- Abul Pharai b. 1226 l. af. 1285 H
- Alain b. 1386 H

> If, as is often the case, the death be known but the birth
> unknown, the line begins with dots, and ends full

- Epaminondas d. 363

> When it is faid that a writer flourijhed at or
> about a particular time, a short full line is drawn
> about two-thirds before and one third after that particular time,
> with three dots before and two after it; because, in
general, men are said to flourish much
> nearer the time of their death than the time of their
birth

- Euclid fl. 300 BC. M

What is left unspecified in the description is the mapping between dots and years?
None of the digital specimens of the Chart of Biography available are high resolution enough to make this out.
Dots seem a little less than decades?

> If it be uncertain even in what
> century a perfon lived or flourished, there is no full
> line made at all, but only dots or broken lines where
> I imagine it is most probable he might have flourished.

Priestley provides no other details about how long this line may be. However,
the examples on this [chart](http://www.enoughroomforspace.org/wp-content/uploads/cartographies-of-time_20130505_0006.jpg) have between six and seven dots, which seem
to correspond to 60--70 years.

Suidas fl. ab. 11 Cent. Cr                  
Hesychius fl. ab. 300 Cr

text                               start2   start1   end1     end2
---------------------------------  -------- -------- -------- --------
Euclid fl. 300 BC. M               -320     -350     -290     -270
Epaminondas d. 363                 303      333       363
Herodotus b. 484 BC. d. af. 413 H           -484      413     423
Suidas fl. ab. 11 Cent. Cr         1000                       1080
Hesychius fl. ab. 300 Cr           250                        330
Agis 241. BC ab. 22                241
Thucydides d. ab. 391. ab. 80 H             311       391

## Miscellaneous

Date abbreviations

------- ----------------------------
d.      died
b.      born
fl.     flourished
af.     after
ab.     about
fl. ab. Considerable uncertainty
B C.    before Christ
------- ----------------------------

## References

-   Priestley, Joseph. 1764 and 1778. Description of a Chart of Biography
    (list of the c2,000 names in the chart). 1st and 7th edn. J. Johnson.
    London. Corrected OCR and annotated by Davis, Stephen Boyd (2009).
    [URL](https://drive.google.com/file/d/0B4KIGf4GncycZGRmNWY4Y2QtZjNjNS00OGEzLWE0MjctMzY0NzFhM2I2YjFj/view?usp=sharing&authkey=CPGfreEB)

-   Davis, Stephen Boyd. "History on the Line: Time as Dimension." Design -
    Issues 28, no. 4 (2012): 4-17. [URL](http://www.jstor.org/stable/23273847).

-   Priestley, Joseph. "A Description of a Chart of Biography with a Catalogue
    of All the Names and the Dates Annexed to Them." 7th edition. 1778.
    Corrected OCR and annotated by Davis, Stephen Boyd. [URL](https://drive.google.com/open?id=0B4KIGf4GncycMmU1ZTdjMGYtM2RiMC00NTE3LThiMTUtZjk5YzkxOThlNGNk).


-   Priestley, Joseph. *A description of a new chart of history with a a Catalogue of All the Names and the Dates Annexed to Them*

    -   2nd edn. 1765. <https://books.google.com/books?id=JHJZAAAAcAAJ>
    -   3rd ed. 1767. <https://books.google.com/books?id=RYZcAAAAcAAJ>
    -   5th ed. 1781. <https://books.google.com/books?id=eCwIAAAAQAAJ>
    -   6th ed. 1777. <https://books.google.com/books?id=4QJcAAAAQAAJ>
    -   1790. <https://books.google.com/books?id=-QJcAAAAQAAJ>
    -   1767. <https://books.google.com/books?id=w5QBAAAAQAAJ>
    -   1805. <https://books.google.com/books?id=6LuuDdFMKQAC>
