---
title: "Joseph Priestly's \"A Chart of Biography\""
author: "Jeffrey B. Arnold"
date: "2/5/2018"
output: html_document
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Describe data


```{r fig.height=2000, fig.width=3000}
library("tidyverse")
library("plotly")
library("priestley")

xaxis <- list(title = "",
              tickmode = "array",
              tickvals = seq(-1200, 1800, by = 100),
              mirror = "all"
              )
xaxis$ticktext <- map_chr(xaxis$tickvals, ~ if_else(.x == 0, 
                                                    str_c("O CE"),
                                                    as.character(abs(.x))))

priestley_bios_1764 %>% 
  ungroup() %>%
  mutate(y = runif(n = length(text)) + as.numeric(as.factor(division))) %>%
  plot_ly() %>%
  add_segments(y = ~ y, yend = ~ y, 
               x = ~ start_2, xend = ~ end_2,
               color = ~ division, text = ~ text, hoverinfo = "text") %>%
  layout(
    xaxis = xaxis,
    yaxis = list(title = "",
                 showticklabels = FALSE,
                 showgrid = FALSE),
    margin = list(b = 120)
  )


```

```{r}
priestley_kings %>% 
  arrange(year) %>%
  mutate(.id = row_number()) %>%
  plot_ly(y = ~ .id, x = ~ year, 
          text = ~ name,
          hovertext = 
          mode = "marker+text")


```


## Descriptions

Examples in which the date of death and the age at death are known. These are drawn with a single solid black line since there is no uncertainty.

> Pindar d. 435 BC. 86 P
> Sophocles d. 406 BC 91 P
> Xenophon d. 359 BC. ab. 90 H
> Plato d. 348 BC. 81 HP
> Terence d. 159 BC. 35 P

> IT is an imperfeftion which must neceiTarily attend every Chart of this 
> nature, that the time of the death, but more efpecially the time of the birth of eminent
> men cannot always be found. In this cafe the  piler muft content hirafelf    with placing his line as
> near as he can conjefture from hiftory where his true
place was, leaving marks to exprefs the uncertainty
> there is attending it. The method I have ufed in
> this Chart is to exprefs certainty by a full line, and
> what is uncertain by dots or a broken line, dispro
> of the dots in the following manner, according to the
> kind or degree of the uncertainty they have to express.


> If it be thought that a person was born or died but
> a little before or after a certain time, the full line
> begins or ends at the certain time, and a single dot only
> is placed at the beginning or end of the line

- Herodotus b. 484 BC. d. af. 413 H

> If history informs us that he was born or died
> about a certain time, a dot is placed just under the
> beginning or end of the full line

- Agis b. 24. BC. ab. 22
- Thucydides d. ab. 391. ab. 80 H

> If the birth be certain but the death uncertain, the line begins full but ends in dots.

Examples: Abul Pharai, Alain

- Abul Pharai b. 1226 l. af. 1285 H
- Alain b. 1386 H

> If, as is often the case, the death be known but the birth
> unknown, the line begins with dots, and ends full

- Epaminondas d. 363

> When it is faid that a writer flourijhed at or
> about a particular time, a short full line is drawn
> about two-thirds before and one third after that particular time,
> with three dots before and two after it; because, in
general, men are said to flourish much
> nearer the time of their death than the time of their
birth

- Euclid fl. 300 BC. M

What is left unspecified in the description is the mapping between dots and years?
None of the digital specimens of the Chart of Biography available are high resolution enough to make this out.
Dots seem a little less than decades?

> If it be uncertain even in what
> century a perfon lived or flourished, there is no full
> line made at all, but only dots or broken lines where
> I imagine it is most probable he might have flourished.

No other details about how long this line may be. However, the examples on the [chart](http://www.enoughroomforspace.org/wp-content/uploads/cartographies-of-time_20130505_0006.jpg) have between six and seven dots.

Examples:

- Suidas fl. ab. 11 Cent. Cr
- Hesychius fl. ab. 300 Cr


text                               start2   start1   end1
----------------------             -------- -------- -------- end2
---------
Euclid fl. 300 BC. M               -320     -350     -290     -270
Epaminondas d. 363                 303      333       363
Herodotus b. 484 BC. d. af. 413 H           -484       413     423
Suidas fl. ab. 11 Cent. Cr         1000                        1080
Hesychius fl. ab. 300 Cr           250                         330
Agis 241. BC ab. 22                241
Thucydides d. ab. 391. ab. 80 H

# Divisions and Classes

Priestley divides the persons in his chart into six broad *divisions*, based on their primary profession.

1. Warriors and Statesmen
2. Divines and Metaphysicians
3. Mathematicians and Physicians
4. Artists and Poets
5. Orators and Critics
6. Historians, Antiquarians, and Lawyers

## Warriors and Statesmen

There are no subcategories

## Divines and Metaphysicians

The subclasses are:

- J: "Signifies a JEW, either a PROPHET, RABBI, or HIGH-PRIEST, &c."
- H P: "HEATHER PHILOSOPHERRS, which are further subdivided according to their sects in the following manner"

  - Ion: Ionic
  - Soc: Socratic
  - Cyr: Cyrenaic
  - Meg: Megaric
  - Ac: Academic
  - Per: Peripatetic
  - Sto: Stoic
  - Cyn: Cynic
  - Ital: Italic
  - Scept: Sceptic
  - Ep: Epicurean
  - Eleack: Eleack
  - Eleat. Eleatic

- F: "Signifies CHRISITIAN FATHERS, including those commonly called HERETICS, who will be found in the Chart between those who are usually called orthodoc and the Jews"
- D: "Christian DIVINES, arbitrarily distinguished from the Fathers, by behind made to commence from the time of Mohammed, and including those called SCHOOLMEN"

- Moh. "MOHAMMADAN DOCTORS"
- Met. "METAPHYSICIANS"
- Mor. "MORALILSTS"
- Pol. "POLITICAL WRITERS, such as Grotius, Puffendorf, and even Algernon Sydney"

## 

- M: "MATHEMATICIANS, including NATURALISTS, NATURAL PHILOSOPHERS and ASTRONOMERS"
- Ph. "PHYSICIANS"
- Chy. "CHYMISTS"


# Fourth Column

- P POET
- Pa. PAINTER
- St. STATUARY
- Mu. MUSICIAN
- Pr. PRINTER
- Act. ACTOR
- Eng. ENGRAVER
- Ar. ARCHITECT

# Fifth Column

- Or. ORATOR
- Bell. CRITIC, including all the Miscellaneous writers of Antiquity, and those who are usually called SOPHISTS
- H. Historian
- Trav. Traveller
- Geo. Geographer
- Ch. Chronologer
- Ant. Antiquity
- L. Lawyer

## Miscellaneous

"Engineer" appears as a pr, but not mentioned in the Description.

Date abbreviations

----
d.      died
b.      born
fl.     flourished
af.     after
ab.     about
fl. ab. Considerable uncertainty
B C.    before Christ

## References

- Priestley, Joseph. 1764 and 1778. Description of a Chart of Biography (list of the c2,000 names in the chart). 1st and 7th edn. J. Johnson. London.
Corrected OCR and annotated by Davis, Stephen Boyd (2009). [URL](https://drive.google.com/file/d/0B4KIGf4GncycZGRmNWY4Y2QtZjNjNS00OGEzLWE0MjctMzY0NzFhM2I2YjFj/view?usp=sharing&authkey=CPGfreEB)
- Davis, Stephen Boyd. "History on the Line: Time as Dimension." Design - Issues 28, no. 4 (2012): 4-17. [URL](http://www.jstor.org/stable/23273847).
- Priestley, Joseph. "A Description of a Chart of Biography with a Catalogue of All the Names and the Dates Annexed to Them." 7th edition. 1778. Corrected OCR and annotated by Davis, Stephen Boyd. [URL](https://drive.google.com/open?id=0B4KIGf4GncycMmU1ZTdjMGYtM2RiMC00NTE3LThiMTUtZjk5YzkxOThlNGNk).


- Priestley, Joseph. *A description of a new chart of history with a a Catalogue of All the Names and the Dates Annexed to Them*
    - 2nd edn. 1765. https://books.google.com/books?id=JHJZAAAAcAAJ
    - 3rd ed. 1767. https://books.google.com/books?id=RYZcAAAAcAAJ
    - 5th ed. 1781. https://books.google.com/books?id=eCwIAAAAQAAJ
    - 6th ed. 1777. (London: J. Johnson) https://books.google.com/books?id=4QJcAAAAQAAJ
    - 1790. https://books.google.com/books?id=-QJcAAAAQAAJ
    - 1767. Warrington. https://books.google.com/books?id=w5QBAAAAQAAJ
    - 1805. https://books.google.com/books?id=6LuuDdFMKQAC